---
editor_options: 
  chunk_output_type: console
---

# Examples of simple feature functionality

## Load necessary libraries
```{r}
library(sf)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(mapview)
```

## Load data from the existing folder (/data)

This is a dataset of occurrences of the black-and-orange flycatcher.
```{r}
barfly_pr <- read.csv("data/barfly_pr.csv")
head(barfly_pr)
```

## Convert it to a shapefile 
```{r}
barfly_pr <- st_as_sf(barfly_pr, coords=c("LONGITUDE","LATITUDE"),crs=4326)
barfly_pr # Let's take a look at it
class(barfly_pr)
```

## Basic visualization
```{r}
plot(barfly_pr$geometry)
axis(1)
axis(2)
box()
grid()
```

## Read in polygon data
```{r}
ghats <- st_read("data/Nil_Ana_Pal.shp")
class(ghats)

# if the coordinate system is known and has not been set
st_crs(ghats) <- 4326  #"+proj=longlat +datum=WGS84 +no_defs"

# plotting
plot(ghats[1,], col = 'grey')
plot(ghats, col = 'green') # terribly bright 
```

## Transforming coordinate systems
```{r}
ghats_proj<- st_transform(ghats, 32643)
ghats_proj
```

## Tapping into tidyverse's capabilities for visualization

## Let's use ggplot2() on the nc shapefile 
## Code adapted from Matt-Strimas Mackey

```{r}
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
ggplot(nc) +
  geom_sf(aes(fill = AREA)) +
  scale_fill_viridis_c("Area") +
  ggtitle("Area of counties in North Carolina") +
  theme_bw()
```


# Let's look at dplyr - the gold standard for data manipulation 

## Couple of functions that are very useful for manipulation of simple features:
## The following verbs operate only on the attribute data and leave the geometries untouched:
  
## select() keeps the specified variables, possibly renaming them
## rename() renames a variable and leaves all others unchanged
## filter() returns the rows that match the given conditions
## mutate() adds new variables based on existing variables
## transmute() creates new variables and drops existing variables
## arrange() sorts by the given variables
## slice() selects rows based on row number
## sample_n() samples n features randomly

## For more such examples, I recommend visiting: http://strimas.com/r/tidy-sf/

## Question usually raised at this point: 

## How do I know what projection to use for my data. Fear not as projectionWizard is here to help - http://projectionwizard.org/
## Also highly recommend using: https://spatialreference.org/

## Some more visualization examples
```{r}
library(rnaturalearth)
library(rnaturalearthdata) # Wonderful dataset with boundaries, polygons 
library(rgeos)

world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)

# basic plot
ggplot(data = world) +
  geom_sf() + theme_bw()

# basic plot with legend
ggplot(data = world) +
  geom_sf() +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("World map", subtitle = paste0("(", length(unique(world$name)), " countries)")) + theme_bw()


```

# goal is to produce several maps - either side by side / or on a grid
# below code was sourced from r-spatial.org
```{r}
(gworld <- ggplot(data = world) +
    geom_sf(aes(fill = region_wb)) +
    geom_rect(xmin = -102.15, xmax = -74.12, ymin = 7.65, ymax = 33.97,  # You need the coordinates for this (Unfortunately)
              fill = NA, colour = "black", size = 1.5) +
    scale_fill_viridis_d(option = "plasma") +
    theme(panel.background = element_rect(fill = "azure"),
          panel.border = element_rect(fill = NA)))
```

# Let's centre a map on the Gulf of Mexico
```{r}
(ggulf <- ggplot(data = world) +
    geom_sf(aes(fill = region_wb)) +
    annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
             fontface = "italic", color = "grey22", size = 6) + 
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) + # Note: coord_sf() uses coordinates from the rectangle used earlier
    scale_fill_viridis_d(option = "plasma") +
    theme(legend.position = "none", axis.title.x = element_blank(), 
          axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
          panel.border = element_rect(fill = NA)))

## what if you want to arrange the above two plots in the same grid
library(cowplot)
plot_grid(gworld, ggulf, nrow = 1, rel_widths = c(2.3, 1))

# To save the plots: 
# ggsave("grid.pdf", width = 15, height =  5)

# also library(patchwork) has very useful functions
```


## Some examples of interactive mapping
```{r}
m1 <- mapview(ghats)
m1

m2 <- mapview(barfly_pr)
m1+m2

# Edit map interactively:
library(mapedit)

a <- mapview(ghats) %>% 
  editMap('ghats')

mapview(a$drawn) + mapview(ghats)

```

